---
output: github_document
---

# `D-SOS`: Dataset shift with outlier scores

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle)
[![License: GPL3](https://img.shields.io/badge/License-GPL3-green.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
[![CRAN](https://www.r-pkg.org/badges/version/dsos)](https://cran.r-project.org/package=dsos)
[![UAI 2022](https://img.shields.io/badge/paper-UAI-yellow)](https://openreview.net/forum?id=S5UG2BLi9xc)
[![Downloads](https://cranlogs.r-pkg.org/badges/dsos)](https://cran.r-project.org/package=dsos)
<!-- badges: end -->

## Overview

`dsos` tests for no adverse shift based on outlier scores. Colloquially,
these tests check whether the new sample is not substantively worse than the old
sample, not if the two are equal as tests of equal distributions do. `dsos`
implements a family of two-sample comparison which assumes that
we have both a training set, the reference distribution, and a test set.

## Installation

If the package is on [CRAN](https://CRAN.R-project.org), install with

```{r cran, eval=FALSE}
install.packages("dsos")
```

From GitHub, install with:

```{r github, eval=FALSE}
# install.packages("remotes")
remotes::install_github("vathymut/dsos")
```

## Example

Use `dsos` to test for adverse shift on the
[`iris`](https://en.wikipedia.org/wiki/Iris_flower_data_set) dataset.
Here, the outlier scores are from extended isolation forest for
density-based out-of-distribution (OOD) detection:

```{r example, eval=TRUE}
library(dsos)
set.seed(12345)
data(iris)
versicolor <- iris[51:100,1:4] # Training sample: Species == 'versicolor'
virginica <- iris[101:150,1:4] # Test sample: Species == 'virginica'
iris_test <- od_pt(x_train = versicolor, x_test = virginica)
plot(iris_test)
```

Among others, `dsos` also implements a method for confidence-based OOD detection
via prediction (resampling) uncertainty.

## Reference

To cite this work and for technical details, please refer to the
[paper](https://openreview.net/forum?id=S5UG2BLi9xc). Sample Bibtex is given below:

```bibtex
@inproceedings{kamulete2022test,
  title     = {Test for non-negligible adverse shifts},
  author    = {Vathy M. Kamulete},
  booktitle = {The 38th Conference on Uncertainty in Artificial Intelligence},
  year      = {2022},
  url       = {https://openreview.net/forum?id=S5UG2BLi9xc}
}
```
